---
title: "FoodExpenditures"
author: "Braden Critchfield and Thomas Olsen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, message=FALSE,warning=FALSE}
library(tidyverse)
library(lmtest)

food <- read_delim("FoodExpenses.txt", delim=" ")
```

# 1: Exploratory Plots and Summary Statistics

```{r}
ggplot(data = food, aes(y = EatingOut, x = Income)) +
  geom_point() +
  ggtitle("Avg. Eating Out Weekly Expenditure by Income Level")

```

There seems to be a positive linear relationship between annual household income and average weekly expenditure on food not cooked at home. However, it looks like the variance might grow larger as Income grows larger.

```{r}
ggplot(data = food, aes(y = EatingOut, x = Income, group=Income)) +
  geom_boxplot() +
  ggtitle("Distributions of Eating Out Weekly Expenditure by Income Level")

```

Because Income is a discrete variable, this shows the distributions of Average weekly eating out expenditure per income level. This still shows a positive relationship between the two variables, and again shows how the variance seems to grow as the income increases.

Here are the summary statistics:

```{r}
summary(food)



```

# 2: Fit homoskedastic linear model

```{r}
food.lm <- lm(EatingOut ~ ., data = food)
summary(food.lm)


```

```{r}
ggplot()+geom_point(mapping=aes(x=fitted(food.lm), y =  resid(food.lm)))

bptest(food.lm)
```

The equal variance assumption is not met. From the residuals vs. fitted values plot, the spread of the data points grows as income grows. This is confirmed by the Breusch-Pagan test, where we can reject the null hypothesis that there is equal variance and assume the variance is unequal.

# 3: Write down heteroskedastic model

# 4: Fit Model from Q3 to Eating Out and check assumptions

```{r}
hetero_sked = gls(data = food, EatingOut ~ Income, weights = 
                    varExp(form=~Income), method="ML")
coef = hetero_sked$coefficients
```

#### Linearity

The avplot shows a linear relationship between eating out and income.

```{r}
avPlots(food.lm)

```

#### Independence

It is a reasonable assumption that the eating out habits of one person does not affect the eating out habits of another.

#### Normality

The below histogram show that the residuals are normally distributed.

```{r}
stdResids_food = resid(object=hetero_sked, type="pearson")


hist(stdResids_food)

```

#### Equal Variance

```{r}
bp_test <- bptest(stdResids_food ~ Income, data = food)
```

The above plot shows that the equal variance has been accounted for by the gls model. The Breusch-Pagan test verifies this with p-value of `r bp_test$p.value` . Since the p-value is above .05 we fail to reject the null hypothesis that there is heteroskedasticity and conclude that that there is no heteroskedasticity.

# 5: Validate predictions via cross-validation

# 6: Report Beta Hat and Variance Parameters

# 7: Test if the economy is not healthy

# 8: Predict your own restaurant spend
