---
title: "FoodExpenditures"
author: "Braden Critchfield and Thomas Olsen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(tidyverse)

food <- read_delim("FoodExpenses.txt", delim=" ")
```


# 1: Exploratory Plots and Summary Statistics
```{r}
ggplot(data = food, aes(y = EatingOut, x = Income)) +
  geom_point() +
  ggtitle("Avg. Eating Out Weekly Expenditure by Income Level")

```

There seems to be a positive linear relationship between annual household income and average weekly expenditure on food not cooked at home. However, it looks like the variance grows larger as Income grows larger.


```{r}
ggplot(data = food, aes(y = EatingOut, x = Income, group=Income)) +
  geom_boxplot() +
  ggtitle("Distributions of Eating Out Weekly Expenditure by Income Level")

```

Because Income is a discrete variable, this shows the distributions of Average weekly eating out expenditure per income level. This still shows a positive relationship between the two variables, and again shows how the variance grows as the income increases.

Here are the summary statistics:
```{r}
summary(food)



```


# 2: Fit homoskedastic linear model



# 3: Write down heteroskedastic model



# 4: Fit Model from Q3 to Eating Out and check assumptions




# 5: Validate predictions via cross-validation




# 6: Report Beta Hat and Variance Parameters




# 7: Test if the economy is not healthy




# 8: Predict your own restaurant spend



